A deeper look into bug fixes: Patterns, replacements, deletions, and additions

### 2016-02-18 MSRChallenge2016.tex

\maketitle
\begin{abstract}

With the rise of automated program repair in the last couple of years, there are a lot of questions to remain unanswered. We performed a study on the Github dump of September 2015 in which we analyzed all the fixing revisions of this dataset, and found all the replacements, additions and deletions made to these files from real life projects in Github.

We found valuable information to guide automatic software repair, such as the most common and least common replacements made by human developers in order to build a successful patch; we also analyzed what are the most likely statements to replace others.

This information will help automatic program repair tools to be more likely to find a patch and also to be more maintainable and user friendly, taking it one step closer to what humans developers do to repair their code.

\end{abstract}

@@

Automatic bug repair is the branch of computer science that deals with automated
ways to repair errors in software. There have been several approaches taken
towards improving the different methodologies to repair bugs in software
automatically~\cite{kim2013,weimer2009,legoues2012,pan2009}. A well known approach is PAR~\cite{kim2013}, which creates 10 different repair templates and applies them to the buggy code in an effort to repair
it. In this paper we have taken 8 out of the 10 PAR templates and tested how
common they are in the repairs made by programmers in the latest official data
dump of Github as of September 2015 provided by~\cite{dyer2013}.

Some other well known approaches~\cite{weimer2009,legoues2012} have been based on the idea of applying a number of mutation operators to the buggy code in order to find a patch. In this opportunity, we study the degree to which these mutation operators apply in practice, taking as a subject of study several million human-made bug fixes from Github. This way it will make it more likely for automatic error repair approaches to succeed in
finding a patch for a particular error, and also, to provide automatic error
repair software with guidelines to make the patches more human-like and
therefore more readable and maintainable by human developers, as studied empirically by~\cite{kim2013}.

We take as reference the study performed by Dongsun Kim et al.~\cite{kim2013} in which they look for the most common ways in which programmers patch bugs in software. The
researchers developed a variation of the tool Genprog~\cite{weimer2009,legoues2012}
with several different templates resembling patterns programmers use to patch
bugs, and then performed an empirical study to evaluate which patches do human programmers prefer.

Since automatic program repair has been well studied in the programming language C. We are now focusing on the programming language Java, and analyzing the behavior of programmers in this particular language.

A good complement for this paper is~\cite{zhong2015}, where they look for additions and
deletions in a smaller data set. In this paper, we focus more on the replacements,
which have been left out of the analysis of~\cite{zhong2015}, and expand their results of the deletions and additions in a broader dataset~\cite{dyer2013}.

\section{Methodology}
Our methodology is applied to the latest official data dump of Github as of
September 2015 provided by~\cite{dyer2013}. In the following subsections, we first
describe data source and objective of our study. We then describe how we detect
bug fixing patterns in PAR to find out how common they are the dataset. Last but
not least, we describe our approach to detect Delete, Replace, and Append
bug-fixing patterns.

\subsection{Data Source and Research Objective}
We used the Boa platform~\cite{dyer2013} to be able to query the Github repository
looking for patterns that humans use to patch errors in the their code. Our main
goal is to provide guidelines for automatic error fixing approaches by providing
necessary information regarding historical replace, delete and append statements
in human written patches.

\subsection{Detecting PAR Bug Fixing Pattern}\label{sec:method}
The bug fixing patterns that we analyze in this work are adopted from Kim et
al. paper~\cite{kim2013}. The following is the description of each considered
pattern. 
\todo{Ferdian: why are we only studying 7?  It's OK to not study all of them,
  but we need to explain why.}
  \begin{enumerate}
	\item {\bf Altering method parameters (AMP)}\\
	{\bf Example:} obj.method(v1,v2) $\rightarrow$ obj.method(v1,v3)\\
	{\bf Description:} This pattern changes the input for method parameters.
	\item {\bf Calling another method with the same parameters (MSM)}\\
	{\bf Example:} obj.method1(param) $\rightarrow$ obj.method2(param)\\
	{\bf Description:} This pattern changes the method name.
	\item {\bf Calling another overloaded method with one more parameter (COM)}\\
	{\bf Example:} obj.method(v1) $\rightarrow$ obj.method(v1,v2)\\
	{\bf Description:} This pattern adds one more parameter to the method.
	\item {\bf Changing a branch condition (CBC)}\\
	{\bf Example:} if(a != b) $\rightarrow$ if(a != b \&\& c == 0)\\
	{\bf Description:} This pattern adds or removes condition.
	\item {\bf Initializing an object (IAO)}\\
	{\bf Example:} Type obj $\rightarrow$ Type obj = new Type()\\
	{\bf Description:} This pattern adds a initialization to object declaration.
	\item {\bf Adding a null checker (ANC)}\\
	{\bf Example:} obj.m1() $\rightarrow$ if(obj!=null){obj.m1()}\\
	{\bf Description:} This pattern inserts a condition to check whether an object is null or not in order to prevent unexpected state.
	\item {\bf Adding an array out of bound checker (AOB)}\\
	{\bf Example:} arr[idx]=0 $\rightarrow$ if(idx<arr.length){arr[idx]=0}\\
	{\bf Description:} This pattern inserts a condition to check that an array index is within bound right before the index is used.
	\item {\bf Adding a collection out of bound checker (COB)}\\
	{\bf Example:} x=col.get(idx) $\rightarrow$ if(idx<col.size()){x=col.get(idx)}\\
	{\bf Description:} This pattern inserts a condition to check that a collection index is within bound before it is used.
\end{enumerate}

The above patterns only cover 8 out of 10 bug fixing patterns in PAR. The other two patterns are: (1) checking that an object is an instance of a target class before casting it; and (2) replacing expression in an if condition by another expression in the same scope. Pattern (1) cannot be analyzed because Boa does not support {\em instanceof} expression. Similarly, pattern (2) requires us to track scope between pre-fix and post-fix version, which is not supported by Boa. Implementing them ourselves might be possible, but it might require too much effort than necessary. Investigating 8 out of 10 patterns should give us a good idea on how common the patterns actually are.

As mentioned before, we analyze bug fixing patterns by utilizing Boa's language
infrastructure~\cite{dyer2013}. It contains a large number of dataset and provide
a domain specific language to perform analysis on source code. Despite this
advantage, Boa still have limited capability that prevents us from performing
precise detection of bug fixing patterns. For example, it is unable to perform a
diff between two file versions, which is needed to compare the pre-fix and the
post-fix version of a buggy file. Thus, rather than finding the exact count of
bug fixing patterns, we will find an approximate count of those patterns. The
following are descriptions on how we detect each pattern.

\begin{enumerate}
\item {\bf AMP Pattern}\\
  To detect AMP pattern, for both pre-fix and post-fix version of a buggy file,
  we create a custom method call signature. This signature contains method name,
  literal parameter, and variable parameter. we consider parameter containing
  custom expression as OTHER. This is due to inexistence of function that can
  print Abstract Syntax Tree (AST) back to source code in Boa. We discard method
  signatures that appear both in pre-fix and post-fix version. We then find
  whether there exists method with the same name and the same number of
  parameters, but different parameter signatures in pre-fix and post-fix
  version. If it exists, we consider that AMP pattern is found.

\item {\bf MSM Pattern}\\
  To detect MSM pattern, we create method signature similar like when detecting
  AMP pattern. This time, we find whether there exists method with the exact
  same parameter signature but different name in pre-fix and post-fix
  version. If it exists, we consider that MSM pattern is found.

\item {\bf COM Pattern}\\
  To detect COM pattern, we create method signature similar like when detecting
  AMP and COM pattern. This time, we find whether there are exists method with
  the same name, but the number of parameters in pre-fix and post-fix version
  differs by one. If it exists, we consider that COM pattern is found.

\item {\bf CBC Pattern}\\
  To detect CBC pattern, we count the number of { \em logical and} and {\em
    logical or} inside if conditional expression for both pre-fix and post-fix
  version. we assume that the addition/removal of logical operators indicate
  addition/removal of condition. If there are count differences between pre-fix
  and post-fix version, we consider that CBC pattern is found.

\item {\bf IAO Pattern}\\
  To detect IAO pattern, we count the number of NEW expression in variable
  declaration for both pre-fix and post-fix version. If there are count
  differences between pre-fix and post-fix version, we consider that IAO pattern
  is found.

\item {\bf ANC Pattern}\\
  To detect ANC pattern, we count the number of if conditional expression that
  contains {\em !=null} or {\em ==null} for both pre-fix and post-fix
  version. If there are count differences between pre-fix and post-fix version,
  we consider that ANC pattern is found.

\item {\bf AOB Pattern}\\
  To detect AOB pattern, we count the number of if conditional expression that
  contains {\em expr<var.length} or {\em var.length>expr} for both pre-fix and
  post-fix version. If there are count differences between pre-fix and post-fix
  version, we consider that AOB pattern is found.

\item {\bf COB Pattern}\\
To detect COB pattern, we count the number of if conditional expression that
contains {\em expr<col.size()} or {\em col.size()>expr} for both pre-fix and
post-fix version. If there are count differences between pre-fix and post-fix
version, we consider that COB pattern is found.

\end{enumerate}
\todo{IMPORTANT: Ferdian, where did the results go?  Please add the results of
  the Par study into results somewhere where it fits.}

\texttt{Assert}	&-&7.48&3.76&0.53&8.30&23.05&0.31&20.04&4.90&4.62&1.30&13.50&7.23&0.03&4.95
\\ \hline
\texttt{Break} &1.00&-&4.08&0.60&9.93&26.03&0.13&25.39&2.48&1.57&1.79&8.39&11.73&0.10&6.77
\\ \hline
\texttt{Continue} &1.74&9.42&-&1.28&11.39&18.25&0.35&22.60&3.80&2.85&2.17&8.98&9.42&0.11&7.63
\\   \hline
\texttt{Do} &0.81&5.26&6.60&0.00&9.44&14.21&0.18&15.86&3.73&1.67&1.97&5.88&6.39&0.03&27.98
\\ \hline
\texttt{For} &0.86&6.28&3.19&0.79&-&22.89&0.09&21.08&5.01&3.34&1.87&10.01&10.71&0.08&13.79
\\ \hline
\texttt{If}&1.64&8.43&2.87&0.60&13.49&-&0.24&26.46&7.45&4.80&2.85&9.89&15.11&0.08&6.11
\\ \hline
\texttt{Label} &1.30&8.33&7.86&1.11&5.18&22.85&-&15.17&3.05&2.04&14.62&10.45&4.16&0.09&3.79
\\ \hline
\texttt{Return} &1.13&9.41&3.11&0.49&13.33&27.24&0.24&-&5.59&3.65&2.55&14.91&12.61&0.12&5.61
\\ \hline
\texttt{Case} &0.78&2.84&2.84&0.39&10.27&31.79&0.16&22.40&-&0.46&2.07&7.37&11.69&0.08&6.87
\\ \hline
\texttt{Switch}&1.14&2.72&3.80&0.55&11.07&34.14&0.13&21.86&0.75&-&1.53&8.65&9.02&0.05&4.58
\\ \hline
\texttt{Synch} &0.80&6.57&2.28&0.43&10.21&24.18&0.05&19.77&6.35&2.07&-&9.16&12.16&0.04&5.93
\\ \hline
\texttt{Throw} &2.11&6.57&2.58&0.48&11.87&18.84&0.17&32.28&4.64&3.30&2.74&-&10.08&0.07&4.27
\\ \hline
\texttt{Try}&0.71&7.41&3.02&0.66&11.73&27.75&0.11&23.24&5.63&2.65&2.58&8.99&-&0.09&5.42
\\ \hline
\texttt{TypeDecl} &0.00&4.51&7.52&1.00&10.28&21.05&0.50&17.79&6.02&1.75&2.01&9.27&11.53&-&6.77
\\ \hline
\texttt{While}& 0.72&8.02&3.82&1.96&23.16&19.78&0.12&16.48&6.56&3.09&1.64&6.81&7.80&0.04&-
			\\ \hline
		\end{tabular}
		}
		\caption{Likelyness to replace the statement in the first column for the statement in the top row, in Java}
\end{table*}

\subsection{Replace Query}
The replace query we designed, is a query in which we create a visitor to traverse all the fixing revisions in this repository~\cite{dyer2013}. Therefore we are analyzing only files
that were already committed and are now being reviewed to fix an error.

We visit each statement in the previous state of that changed
file, before the change was applied and we count the amount of appearances of
the first statement we are analyzing (the one that was deleted), and then we
count the amount of appearances of the second statement we are analyzing (the
one that was inserted). We save these values, and do the same for the latest
version of the file and save these values as well.

We then compare these results to see if either the first and second statements
being analyzed increased or decreased. Depending on these we have two
conditionals:

If the amount of occurrences of the first statement decreased and the amount of
occurrences of the second statement increased on the same file, then we say that
the first statement was replaced by the second statement in that file. Likewise,
if the amount of occurrences of the first statement increased and the amount of
occurrences of the second statement decreased on the same file, we say that the
second statement was replaced by the first statement in that file. We count the
amount of files in which this happens for each of these two cases.

\section{Results}

% (This is the old version)
% This section investigates how do developer fix bugs from three different
% perspectives. To begin with, we show some characteristics of real-world bug
% fixes.  Afterwards, we look deeper to find out what kinds of statements are
% usually added, modified or deleted by human developers to fix bugs. Lastly, we
% show how common PAR's fixing templates are in our dataset.

This section shows our analysis results. To begin with, we discuss the dataset
we are using, and some characteristics of real-world bug fixes. Afterwards, we
provide insights for two main categories of automatic program repair approaches.
On the one hand, some approaches rely on mutation operators to
modify the source code~\cite{legoues2012,weimer2009}, but we have limited knowledge about what kinds of
statements are usually modified. Our analysis results will fill in this
knowledge gap. On the other hand, some automatic program repair approaches~\cite{kim2013} search for specific patterns and modify source code according to
predefined templates. Our results provide evidence of how common those
patterns are in practice.

\subsection{Dataset and Characteristics}

We did our analysis on the \emph{2015 September/Github} dataset on Boa. This
dataset includes $7,830,023$ Java projects with $23,229,406$ revisions. Among
all the revisions, $4,590,679$ revisions are identified by Boa as bug fixing
revisions. We hope that this population is big enough to reflect the bug fixing
practice.

The rest of this subsection shows some characteristics of real-world bug fixes.
To begin with, we are interested in:

\begin{quote}
	How many files are changed to fix a bug?
\end{quote}

To answer this question, we identify all the bug fixing revisions in our dataset
and count the number of changed files. A file was changed if (1) it was a new
file; (2) or it was modified; (3) or it was deleted. In total, $52,052,571$
files got changed. Thus, on average each bug fixing revision changes 11.3 files.
This number is surprisingly high because most automatic program repair
techniques assume that bugs are local in most cases. There could be several
reasons behind this number. It could be the case that other software artifacts
(e.g.  documentation) are updated after bug fixing. It could also be the case
that bug fixing commit also contains changes that are not related to fixing bug
(e.g. refactoring, adding new features by creating new files). To better
understand why so many files are changed, we continue to ask:

\begin{quote}
	What are the file types of those changed files?
\end{quote}

In Boa, each changed file is represented by a \emph{ChangedFile} node, and it
has a field called \emph{FileKind}, describing the type of each file. As
mentioned earlier, a \emph{ChangedFile} could be either a new file or an
existing file (if the bug fixing revision modified or deleted the file).
Table~\ref{tbl:fileType} shows the file types that Boa could identify. To answer
this question, we iterate all the changed files in each bug fixing revision and
count the total number for each file type. Results are shown in
Table~\ref{tbl:fileType}.

@@

\begin{quote}
	How often do developers introduce new classes, methods, fields, variables in
	bug fixing commits?
\end{quote}

This research question is interesting because most automatic program repair
approaches often do not create new classes/methods/fields/variables to fix bugs.
Most approaches guess and synthesize the patches based on existing source code.

@@

\subsection{What changes are common in real-world bug fixes?}

Most automatic program repair approaches mutate source code using predefined
mutation operators, which represent different ways of modifying source code. In
order to generate patches that make more sense to human developers, it would be
beneficial to first investigate how developers mutate source code to fix bugs.

@@

\subsubsection{Likeliness of replacement}
This study is directly applicable to automatic software repair. Most common
tools for automatic software repair~\cite{kim2013,weimer2009,legoues2012}
first find a faulty statement using several different fault localization
techniques~\cite{fry2010}, and once they have that faulty statement, they apply
different kinds of edits to it, usually at random. Which makes it equally likely
to replace an statement for another without looking at the kinds of statements
being replaced.

>>>>

\maketitle
\begin{abstract}
  Many implementations of recent work to automatically repair bugs in software
  target programs written in C. Work that targets Java often begins from, or at
  least compares to, direct translations of such techniques to a Java context.
  However, Java and C are very different languages, and these differences should be
  studied and understood if existing approaches are to be profitably translated
  from one to another.  We conducted a large-scale study of bug-fixing commits
  to Java projects, focusing on assumptions underlying common
  search-based repair approaches.  We make a number of observations that can be
  leveraged to guide automatic software repair to target Java specifically,
  including common and uncommon statement replacements as conducted by human
  developers and the applicability of previously-proposed patch construction
  operators in the Java context.  We anticipate that these observations can
  inform more effective patch construction techniques for Java, both in raw
  numbers of bugs fixed and the usability and maintainability of the resulting
  patches. 
\end{abstract}

@@

Research in automatic bug repair seeks to develop techniques that can
automatically construct patches to repair software bugs, which are prevalent in
practice. There has been considerable attention paid to this problem in recent
years~\cite{kim2013,weimer2009,legoues2012,pan2009,so,many,more}.
Mutation-based, or generate-and-validate approaches, typically operate by
generating a large number of candidate patches using a pre-defined set of
mutation operators, and then validating their correctness with respect to a set
of test cases.  These mutation operators range from simple, coarse granularity
statement-level mutations~\cite{examples} to human-constructed templates learned
from a large corpus of previous human bug-fixing commits~\cite{kim2013}.

A considerable proportion of this previous work is implemented to target
programs written in C, with only limited attention paid to programs in other
languages, like Java~\cite{kim2013}.  At least partially as a result,
researchers targeting Java must either start from or at least compare against
imperfect baselines, that is, Java-based implementations of techniques
originally implemented for C~\cite{nopol}. %  Even PAR~\cite{kim2013}, perhaps the
% best-known previously-proposed Java-based program repair technique, both starts
% from and is directly compared to a Java implementation of GenProg~\cite{genprog}, which
% targets the C programming language.  
This C focus extends to empirical studies
of bug fixing practice, used to inform research questions in automatic
repair~\cite{zhong2015}.

This is an oversight, because, to put it simply: Java and C are very different
languages.  Consider GenProg~\cite{genprog,cites}, a well-known technique in
this space: GenProg uses a genetic programming heuristic to combine
statement-level mutations into patches to address a particular defect.
``Statement'' is a well-defined semantic unit in the C programming
language~\cite{cspec}; to which semantic unit in Java does this concept map?
Can we learn more
intelligent rules to inform candidate modifications
(cf. Prophet~\cite{rinardPopl2016}, for C)?  What types of
semantic scoping rules should be used to limit mutations, given the relative
stringence of the Java compilers?

In this paper, we study bug-fixing commits to Java programs, taking as a subject
of study several million human-made bug fixes from Github. We study broad
characteristics of changes to fix bugs; the applicability of previously-proposed
templates to repairing defects in Java programs~\cite{kim2013}; and the nature
of additions, replacements, of statements in bug-fixing commits. We make
observations to directly guide future research in automatic repair to Java
programs, to increase the success rate of such techniques and the degree to
which the patches are human-like and therefore more readable and maintainable by
human developers; this is important, because patch quality is an important
concern in this research area~\cite{Qi2013}.

% On the other hand, some approaches~\cite{kim2013} search for specific patterns
% and modify source code according to predefined templates. A well known approach
% is PAR~\cite{kim2013}, which creates 10 different repair templates and applies
% them to the buggy code in an effort to repair it. In this paper we have taken 8
% out of the 10 PAR templates and tested how common they are in the repairs made
% by programmers in the latest official data dump of Github as of September 2015
% provided by~\cite{dyer2013}.  Our results provide evidence of how common those
% patterns are in practice. We take as reference the study performed by Dongsun
% Kim et al.~\cite{kim2013} in which they look for the most common ways in which
% programmers patch bugs in software. The researchers developed a variation of the
% tool Genprog~\cite{weimer2009,legoues2012} with several different templates
% resembling patterns programmers use to patch bugs, and then performed an
% empirical study to evaluate which patches do human programmers prefer.

A good complement for this paper is~\cite{zhong2015}, where they look for
additions and deletions in a smaller data set. In this paper, we focus more on
the replacements, which have been left out of the analysis of~\cite{zhong2015},
and expand their results of the deletions and additions in a broader
dataset~\cite{dyer2013}.
\todo{note to self (CLG) should mention the work from Earl, Yuriy, and others
  that looked at the redundancy assumption.}

\todo{roadmap}

\section{Dataset and Characteristics}

We begin by broadly characterizing the bug-fixing changes to Java code in our
dataset. 
We perform our analysis on the \emph{2015 September/Github} dataset on Boa, 
which includes $7,830,023$ Java projects with $23,229,406$ revisions. Among
all the revisions, $4,590,679$ revisions are identified by Boa as bug fixing
revisions.

\paragraph{How many files are changed to fix a bug?}
To answer this question, we identify all the bug fixing revisions in our dataset
and count the number of changed files. A file was changed if (1) it was a new
file; (2) or it was modified; (3) or it was deleted. In total, $52,052,571$
files got changed. Thus, on average each bug fixing revision changes 11.3 files.
This number is surprisingly high because most automatic program repair
techniques assume that bugs are local in most cases. There could be several
reasons behind this number. It could be the case that other software artifacts
(e.g.  documentation) are updated after bug fixing. It could also be the case
that bug fixing commit also contains changes that are not related to fixing bug
(e.g. refactoring, adding new features by creating new files). To better
understand why so many files are changed, we continue to ask:

\paragraph{What are the file types of those changed files?}
In Boa, each changed file is represented by a \emph{ChangedFile} node, and it
has a field called \emph{FileKind}, describing the type of each file. As
mentioned earlier, a \emph{ChangedFile} could be either a new file or an
existing file (if the bug fixing revision modified or deleted the file).
Table~\ref{tbl:fileType} shows the file types that Boa could identify. To answer
this question, we iterate all the changed files in each bug fixing revision and
count the total number for each file type. Results are shown in
Table~\ref{tbl:fileType}.

@@

\paragraph{
How often do developers introduce new classes, methods, fields, variables in
	bug fixing commits?}
This research question is interesting because most automatic program repair
approaches often do not create new classes/methods/fields/variables to fix bugs.
Most approaches guess and synthesize the patches based on existing source code,
and direct translations of approaching from C to Java do not naturally consider
object-oriented concerns, like fields or classes.

@@

\section{PAR Bug Fixing Patterns}\label{sec:method}

Par is a well-known technique for patch generation that 
On the other hand, some approaches~\cite{kim2013} searches for specific patterns
and modifies java code according to predefined templates, constructed by humans
to cover a large set of bug fixes from an existing corpus.  As one of the only
techniques to directly target Java, it is an important source of possible
mutations to study to inform future directions. 
We therefore search for and analyze most of the bug fixing patterns proposed by
Kim et al.~\cite{kim2013} and assess their prevalence on this distinct
dataset by using Boa's language infrastructure~\cite{dyer2013} to identify the patterns
in the commits in the dataset.

% A well known approach
% is PAR~\cite{kim2013}, which creates 10 different repair templates and applies
% them to the buggy code in an effort to repair it. In this paper we have taken 8
% out of the 10 PAR templates and tested how common they are in the repairs made
% by programmers in the latest official data dump of Github as of September 2015
% provided by~\cite{dyer2013}.  Our results provide evidence of how common those
% patterns are in practice. We take as reference the study performed by Dongsun
% Kim et al.~\cite{kim2013} in which they look for the most common ways in which
% programmers patch bugs in software. The researchers developed a variation of the
% tool Genprog~\cite{weimer2009,legoues2012} with several different templates
% resembling patterns programmers use to patch bugs, and then performed an
% empirical study to evaluate which patches do human programmers prefer.

\subsection{Detecting PAR patterns in Boa}

Boa's capabilities are powerful, but are limited in the precision it enables in
our detection of these patterns. For example, it is unable to directly perform a
diff between two file versions. Thus, rather than finding the exact count of
bug fixing patterns, we will find an approximate count by processing pre- and
post- fix files separately.  Additionally, there are
two of the originally proposed 10 patterns that cannot be easily detected by
Boa, as we describe below.

The patterns we search for are as follows, including our mechanism for
detecting them using Boa:

\vspace{1ex}
\noindent {\bf Altering method parameters (AMP)}.  This pattern changes the
input for method parameters.  {\bf Example:} obj.method(v1,v2) $\rightarrow$
obj.method(v1,v3). To detect this pattern, for both pre-fix and post-fix version
of a buggy file, we create a custom method call signature. This signature
contains method name, literal parameter, and variable parameter. we consider
parameter containing custom expression as OTHER. This is due to inexistence of
function that can print Abstract Syntax Tree (AST) back to source code in
Boa. We discard method signatures that appear both in pre-fix and post-fix
version. We then find whether there exists method with the same name and the
same number of parameters, but different parameter signatures in pre-fix and
post-fix version. If it exists, we consider that AMP pattern is found.

\vspace{1ex}
\noindent {\bf Calling another method with the same parameters (MSM)}.  This
pattern changes a method name.  {\bf Example:} obj.method1(param) $\rightarrow$
obj.method2(param).  To detect this pattern, we create method signature similar
like when detecting AMP pattern. This time, we find whether there exists method
with the exact same parameter signature but different name in pre-fix and
post-fix version. If it exists, we consider that MSM pattern is found.

\vspace{1ex}
\noindent
{\bf Calling another overloaded method with one more parameter (COM)}. This
pattern adds one more parameter to a method. 
	{\bf Example:} obj.method(v1) $\rightarrow$ obj.method(v1,v2).
  To detect COM pattern, we create method signature similar like when detecting
  AMP and COM pattern. This time, we find whether there are exists method with
  the same name, but the number of parameters in pre-fix and post-fix version
  differs by one. If it exists, we consider that COM pattern is found.

\vspace{1ex}
\noindent
 {\bf Changing a branch condition (CBC)}. This pattern adds or removes a condition.
	{\bf Example:} if(a != b) $\rightarrow$ if(a != b \&\& c == 0).
  To detect CBC pattern, we count the number of { \em logical and} and {\em
    logical or} inside if conditional expression for both pre-fix and post-fix
  version. we assume that the addition/removal of logical operators indicate
  addition/removal of condition. If there are count differences between pre-fix
  and post-fix version, we consider that CBC pattern is found.

\vspace{1ex}
\noindent
 {\bf Initializing an object (IAO)}. This pattern adds a initialization to object declaration.
  To detect IAO pattern, we count the number of NEW expression in variable
  declaration for both pre-fix and post-fix version. If there are count
  differences between pre-fix and post-fix version, we consider that IAO pattern
  is found.
	{\bf Example:} Type obj $\rightarrow$ Type obj = new Type().

\vspace{1ex}
\noindent
 {\bf Adding a null checker (ANC)}.  This pattern inserts a condition to check whether an
    object is null or not in order to prevent unexpected state.
	{\bf Example:} obj.m1() $\rightarrow$ if(obj!=null){obj.m1()}
  To detect ANC pattern, we count the number of if conditional expression that
  contains {\em !=null} or {\em ==null} for both pre-fix and post-fix
  version. If there are count differences between pre-fix and post-fix version,
  we consider that ANC pattern is found.

\vspace{1ex}
\noindent
 {\bf Adding an array out of bound checker (AOB)}. This pattern inserts a condition to check that an array
    index is within bound right before the index is used.
	{\bf Example:} arr[idx]=0 $\rightarrow$ if(idx<arr.length){arr[idx]=0}.
  To detect AOB pattern, we count the number of if conditional expression that
  contains {\em expr<var.length} or {\em var.length>expr} for both pre-fix and
  post-fix version. If there are count differences between pre-fix and post-fix
  version, we consider that AOB pattern is found.

\vspace{1ex}
\noindent
 {\bf Adding a collection out of bound checker (COB)}.  This pattern inserts a condition to check that a
    collection index is within bound before it is used.
	{\bf Example:} x=col.get(idx) $\rightarrow$ if(idx<col.size()){x=col.get(idx)}\\
To detect COB pattern, we count the number of if conditional expression that
contains {\em expr<col.size()} or {\em col.size()>expr} for both pre-fix and
post-fix version. If there are count differences between pre-fix and post-fix
version, we consider that COB pattern is found.

\vspace{1ex} The above patterns only cover 8 out of 10 bug fixing patterns in
PAR. The other two patterns are: (1) checking that an object is an instance of a
target class before casting it; and (2) replacing expression in an if condition
by another expression in the same scope. Pattern (1) cannot be analyzed because
Boa does not support the {\em instanceof} expression. Similarly, pattern (2)
requires us to track scope between pre-fix and post-fix version, which is not
supported by Boa. Implementing them ourselves might be possible, but it might
require too much effort than necessary. We speculate that investigating 8 out of
10 patterns should give us a good idea on how common the patterns actually are.

\subsection{Results} \label{sec:freqfixpattern}

\begin{table}[!htb]
	\caption{Frequency of Bug Fixing Patterns}\label{tab:freqpattern}
	\centering
	\begin{tabular}{lr} 
		\hline
		& \textbf{GitHub}\\
		\hline
		\#Buggy Files & 46,301,429 \\ 
		\#AMP Pattern & 901,083 \\
		\#MSM Pattern & 783,073 \\
		\#COM Pattern & 270,128 \\ 
		\#CBC Pattern & 1,959,377 \\  
		\#IAO Pattern & 1,308,006 \\  
		\#ANC Pattern & 1,340,561 \\  
		\#AOB Pattern & 128,016 \\  
		\#COB Pattern & 262,915 \\   
		\hline
	\end{tabular}
\end{table}
\todo{Ferdian: please add a column that shows the number sin terms of
  percentages, not just raw numbers.}

After running the bug pattern detection approach on Boa's dataset, I obtained
the result shown in Table~\ref{tab:freqpattern}. The result shows that AMC, MSM,
COM, CBC, IAO, ANC, AOB, and COB pattern is found on 1.95\%, 1.69\%, 0.58\%,
4.23\%, 2.82\%, 2.90\%, 0.28\%, and 0.57\% of buggy files. We can see that the
most common pattern among these patterns is CBC and the least common pattern is
AOB. If we assume that these patterns never appear together in the dataset,
these patterns cover 14.78\% of buggy files. Kim et al. mentioned that the
patterns that they use cover almost 30\% of real
patches~\cite{kim2013}. Although we only detect 8 out of 10 bug fixing patterns,
it is hard to believe that the remaining two patterns cover 15\% of buggy files,
especially 14.78\% is a number we get when we assume that the patterns that we
detect never appear together. Moreover, the percentage of each pattern is based
on imprecise approximation which is likely to represent an upper bound rather
than a lower bound. This suggests that common bug fixing patterns might not be
as common as it appeared to be.

\section{Statement-level mutations}

\texttt{Assert}	&-&7.48&3.76&0.53&8.30&23.05&0.31&20.04&4.90&4.62&1.30&13.50&7.23&0.03&4.95
\\ \hline
\texttt{Break} &1.00&-&4.08&0.60&9.93&26.03&0.13&25.39&2.48&1.57&1.79&8.39&11.73&0.10&6.77
\\ \hline
\texttt{Continue} &1.74&9.42&-&1.28&11.39&18.25&0.35&22.60&3.80&2.85&2.17&8.98&9.42&0.11&7.63
\\   \hline
\texttt{Do} &0.81&5.26&6.60&0.00&9.44&14.21&0.18&15.86&3.73&1.67&1.97&5.88&6.39&0.03&27.98
\\ \hline
\texttt{For} &0.86&6.28&3.19&0.79&-&22.89&0.09&21.08&5.01&3.34&1.87&10.01&10.71&0.08&13.79
\\ \hline
\texttt{If}&1.64&8.43&2.87&0.60&13.49&-&0.24&26.46&7.45&4.80&2.85&9.89&15.11&0.08&6.11
\\ \hline
\texttt{Label} &1.30&8.33&7.86&1.11&5.18&22.85&-&15.17&3.05&2.04&14.62&10.45&4.16&0.09&3.79
\\ \hline
\texttt{Return} &1.13&9.41&3.11&0.49&13.33&27.24&0.24&-&5.59&3.65&2.55&14.91&12.61&0.12&5.61
\\ \hline
\texttt{Case} &0.78&2.84&2.84&0.39&10.27&31.79&0.16&22.40&-&0.46&2.07&7.37&11.69&0.08&6.87
\\ \hline
\texttt{Switch}&1.14&2.72&3.80&0.55&11.07&34.14&0.13&21.86&0.75&-&1.53&8.65&9.02&0.05&4.58
\\ \hline
\texttt{Synch} &0.80&6.57&2.28&0.43&10.21&24.18&0.05&19.77&6.35&2.07&-&9.16&12.16&0.04&5.93
\\ \hline
\texttt{Throw} &2.11&6.57&2.58&0.48&11.87&18.84&0.17&32.28&4.64&3.30&2.74&-&10.08&0.07&4.27
\\ \hline
\texttt{Try}&0.71&7.41&3.02&0.66&11.73&27.75&0.11&23.24&5.63&2.65&2.58&8.99&-&0.09&5.42
\\ \hline
\texttt{TypeDecl} &0.00&4.51&7.52&1.00&10.28&21.05&0.50&17.79&6.02&1.75&2.01&9.27&11.53&-&6.77
\\ \hline
\texttt{While}& 0.72&8.02&3.82&1.96&23.16&19.78&0.12&16.48&6.56&3.09&1.64&6.81&7.80&0.04&-
			\\ \hline
		\end{tabular}
		}
		\caption{Likeliness to replace the statement in the first column for the statement in the top row, in Java}
\end{table*}

Most automatic program repair approaches mutate source code using predefined
mutation operators, which represent different ways of modifying source code. In
order to generate patches that make more sense to human developers, it would be
beneficial to first investigate how developers mutate source code to fix bugs. 
Thus, beyond characterizing the degree to which previously-proposed Java-specific
operators apply in our dataset, we further characterized the nature of
coarser-grained replacement operators, which are also commonly translated from C
techniques to Java.  We did this via a query that visits all fixing revisions in
the repository; as a result, we analyze files to which changes have been made
and thuswere already committed and are now being reviewed to fix an error.

\subsection{Replacements}
\todo{The biggest problem right now is the lack of integration between this
  study and Chu Pan's.  Please work together to find a way to unify it.  One of
  you will need to cut some of your text.  I don't care who, but we need to find
  a way to tell one story.}

Because direct diffs are difficult to identify, we heuristically approximate the
extent to which one statement type appears to be ``replaced'' by another across
this dataset.  To do this, for each modified file, we count the number of
appearances of each statement type in the file pre- and post-commit.  We then
compare the results to see how many of each statement type was removed, and how
many inserted, to roughly characterize the types of replacement that happen at a
per-file level.  For each statement type (as tagged by the Boa infrastructure
for Java), we process the results as follows: 
if number occurrences of one statement type decreased post-fix and the number of
another type increased, we say that
the first statement type was replaced by the second statement type for that
file.  We collate these results over all statement types and all files modified
in bug fixing commits in the dataset.

This study is directly applicable to automatic software repair. Most common
tools for automatic software repair~\cite{kim2013,weimer2009,legoues2012}
first find a faulty statement using several different fault localization
techniques~\cite{fry2010}, and once they have that faulty statement, they apply
different kinds of edits to it, usually at random. Which makes it equally likely
to replace an statement for another without looking at the kinds of statements
being replaced.

\subsection{Deletions and insertions}

@@
